Stock enhancement undermines community stability

### 2022-01-27 rmd/manuscript.Rmd
edit: rmd

source(here::here("code/data_fmt_fishdata.R"))

@@

@@

**Affiliations:** ^1^ Department of Biology, University of North Carolina at Greensboro, Greensboro, NC 27412, USA ^2^ Department of Ecology, Evolution and Behavior, University of Minnesota, St. Paul, MN 55108, USA ^3^ National Institute for Environmental Studies, Tsukuba, Ibaraki 305-8506, Japan ^4^ Shiraume Gakuen College, Kodaira, Tokyo 187-8570, Japan

@@

where $r_i$ is the intrinsic growth rate (drawn from a uniform distribution as $r_i \sim Unif(0,2)$ except $r_1$), $\alpha_{ij}$ the competition coefficient of species $j$ on species $i$ [drawn from an exponential distribution as $\alpha_{ij,i \ne j} \sim Exp(1/\bar{\alpha})$; intraspecific competition $\alpha_{ii}=1$], $K_i$ the carrying capacity, $R_t$ the number of released individuals for the enhancement of species 1, and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations that follow an normal distribution with a mean of zero and a standard deviation of $\sigma_{\epsilon}$. The parameter $\phi_i$ controls the relative fitness of captive-bred individuals as follows:

@@

In our simulation, we considered `r nrow(param_set)` simulation scenarios with different values of the following ecological parameters: intrinsic growth of the enhanced species ($r_1=$ `r unique(sim_result$r1)`), average strength of interspecific competition ($\bar{\alpha}=$ `r unique(sim_result$alpha)`), carrying capacity ($K=$ `r unique(sim_result$k)`), relative fitness of captive-bred individuals ($f_R=$ `r unique(sim_result$phi)`), and degree of environmental stochasticity ($\sigma_{\epsilon}=$ `r unique(sim_result$sd_env)`). We chose these parameters because our sensitivity analysis revealed that these parameters were influential in altering the relationship between community dynamics and stock enhancement (**Supporting Information**). Under each scenario with a unique parameter combination, we ran `r n_sim` time steps of `r max(sim_result$n_rep)` independent communities (i.e., simulation replicates) with varying numbers of released individuals $R$ ($0-500$). The last `r unique(sim_result$n_timestep)` time steps were used to obtain the following summary statistics of total community density $\sum_i^SN_i$: the coefficient of variation (CV), the number of species persist (defined as $N_i > 0.01$ at $t=1600$), the temporal mean, and the temporal SD. In addition, temporal means and SDs were calculated for the enhanced ($N_1$) and unenhanced species ($\sum_{i,i \ne 1}^S N_i$) separately to infer mechanisms that underlie the community-level response.

@@

The only scenario where stock enhancement had little influence on community dynamics was when carrying capacity was sufficiently large ($K=500$; **Figures S7-S12**). In particular, stock enhancement increased the average density of the enhanced species with low population growth rate ($r_1=0.5$; **Figures S7 and S8**). This result may explain why some of the best evidence for successful stock enhancement comes from endangered species that typically suffer from increased mortality [@johnsonGeneticRestorationFlorida2010; @oppelPopulationReinforcementDemographic2021]. More importantly, the contrasting community response at different levels of carrying capacity provides deeper insights into ecosystem management. One of the core motivations for stock enhancement is to mitigate or reverse declining trends of biodiversity due to human impacts, such as habitat loss and fragmentation [@chaseEcosystemDecayExacerbates2020]. However, our results suggest that stock enhancement will not bring desired ecological outcomes in degraded habitats because human influences compromised environmental capacity.

@@

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of selective stock enhancement of masu salmon *Oncorhynchus masou masou* (**Figure 1E**) on the long-term stability of stream fish communities. Masu salmon is one of the most valuable freshwater resources in Hokkaido Island, Japan, and extensive hatchery programs have been implemented [@kitadaLessonsJapanMarine2020]. In protected watersheds where any forms of exploitation are prohibited (including recreational angling) (**Figure 1C and 1D**), long-term monitoring has been conducted over the past two decades along with official records of stock enhancement [@teruiMetapopulationStabilityBranching2018]. Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. In this study, we used the long-term data (1999-2019) of stream fish communities at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` protected watersheds (see Methods for selection criterion). Using hierarchical Bayesian models, we quantified the effect of stock enhancement on community dynamics while accounting for potential confounding effects of climates and local abiotic factors.

As our theory predicts, stream fish communities showed greater temporal fluctuations (higher CV) in watersheds with greater numbers of fish released (**Figure 1B**). The positive relationship between the CV and the number of fish released was strongly supported by the statistical model (**Figure 2**), in which the probability of the regression coefficient being positive was `r op(df_beta$all$prob_positive[df_beta$all$response == "cv"])` (**Table S7**). This pattern was associated with the reduced long-term average of total community density, and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Tables S7-S9** and **Figure 1B and 2**). In the meantime, the SDs had vague relationships with stock enhancement, and neither positive or negative relationship was supported by statistical analysis (**Figure 1B and 2**). At the population level, similar responses to stock enhancement have been reported in other fish and aquatic invertebrates [@arakiHatcheryStockingHelp2010]. Our analysis builds upon these findings by illustrating the community-level response to stock enhancement in nature.

@@

Like any non-experimental studies, we cannot exclude the possibility that our results were driven by spurious correlations. However, in the protected watersheds, any exploitation of aquatic species and alteration of in-stream habitats (e.g., construction of river-crossing structure) are strictly regulated; therefore, it is difficult to envision that unmeasured human influences played a major role in the observed relationships. Further, we have statistically controlled natural variation in environmental factors, including climatic and land use variables (see **Methods** and **Tables S7-S9**). This unique study setup may have helped uncover the close agreement between theory and empirical patterns.

@@

We employed a multispecies Ricker model. In the basic formula without stock enhancement, population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient between species $i$ and $j$, $K_i$ the carrying capacity, and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations that obey an normal distribution $Normal(0, \sigma_{\epsilon}^2)$. We modified this formula to include the effects of stock enhancement on reproduction and competition as follows:

@@

In this model, the intrinsic growth rates for unenhanced species $r_{i,i \ne 1}$ and interspecific competition $\alpha_{ij}$ are random draws from a uniform ($r_{i,i \ne 1} \sim Unif(0,r_{max})$) and an exponential distribution ($\alpha_{ij} \sim Exp(1/\bar{\alpha})$), respectively. We assumed constant values of intraspecific competition ($\alpha_{ii}=1$), carrying capacity $(K_i = K)$ and the number of release $(R_t=R)$.

Prior to the main simulation analysis, we performed an extensive sensitivity analysis to identify parameters that strongly influence the relationship between community dynamics and stock enhancement. We identified five influential parameters, of which multiple values were considered in the main simulation as follows: carrying capacity ($K=$ `r unique(param_set$k)`), intrinsic growth of enhanced species ($r_1 =$ `r unique(param_set$r1)`), degree of environmental stochasticity ($\sigma_{\epsilon}=$ `r unique(param_set$sd_env)`), relative fitness of captive-bred individuals ($f_R=$ `r unique(param_set$phi)`), and average strength of interspecific competition ($\bar{\alpha}=$ `r unique(param_set$alpha)`). We fixed values of the following parameters: the number of species ($S=$ `r unique(param_set$n_species)`) and the maximum intrinsic growth rate of unenhanced species ($r_{max}=$ `r unique(param_set$r_max)`). This simulation setup resulted in `r nrow(param_set)` sets of parameter combinations that cover a range of ecological scenarios.

Under each parameter combination, we ran `r n_sim` time steps of `r max(sim_result$n_rep)` independent communities (i.e., simulation replicates). The number of released individuals $R$ was drawn randomly from a uniform distribution for each simulation replicate as $R \sim Unif(0,500)$. We initialized the community with populations of each species drawn from a poisson distribution with a mean of five. We repeated the seeding procedure every 10 time steps over the first `r unique(sim_result$n_warmup)` time steps to allow species to establish populations and reach an equilibrium without stock enhancement (initialization period). After the initialization period, we released $R$ individuals of the enhanced species every time step over the next `r unique(sim_result$n_burnin)` time steps to reach a new equilibrium with selective stock enhancement (burn-in period). After the burn-in period, we continued the simulation run with stock enhancement and saved the last `r unique(sim_result$n_timestep)` time steps to obtain the coefficient of variation, temporal mean, and SD of whole community density ($\sum_i^SN_j$). Temporal means and SDs were also calculated for the enhanced species ($N_1$) and unenhanced species ($\sum_{i,i \ne 1}^S N_i$) separately. We summarized values of simulation parameters in **Table S1**.

@@

**Time-series data.** We assembled time-series data of fish abundance across `r n_distinct(d0$site_id)` sites within `r n_distinct(d0$river)` protected watersheds of Hokkaido Island, Japan. At these sites, the long-term monitoring program began in 1971 but effective sampling methods (a combination of electrofishing and cast net sampling) were implemented in 1999. Most data were collected in summer to fall with irregular sampling intervals (0 to 3 year intervals for most cases). We confined our analysis to the sites that meets the following selection criteria: (1) the observation span (from the first to the last year of observation) exceeds 10 years, (2) the number of observation years exceeds five years, and (3) masu salmon are observed at least twice during the observation period. As a result, we used time-series data of fish abundance at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` during a period of 1999-2019.

**Fish stocking.** At the protected watersheds, stocking of masu salmon began in 1950s. A mixture of captive-bred and wild-born individuals were released in spring (fry [age 0+] and smolt [age 1+] stages) and fall (juvenile stage [age 0+]). For each watershed, we examined historical records and assembled annual data of stock enhancement (the number of individuals released) for a period of 1999-2019. Although fish were released at multiple locations within a watershed, we were unable to extract stocking data at the site level.

**Environmental data.** At each sampling site, we collected the following data as potential covariates: upstream watershed area (km^2^; a proxy for stream size), proportional land use in the upstream watershed (forest), local climates (annual mean air temperature [unit] and cumulative precipitation [mm]), and ocean productivity (Chlorophyll a [unit]). Annual data of ocean productivity (2003-2019) were extracted from OceanColor (<https://oceancolor.gsfc.nasa.gov/>) and the averaged value within the 30-km radius of each river mouth was calculated.

@@

Our goal is to compare temporal community dynamics across sites. However, the data are not directly comparable because of observation errors (e.g., different observers) and missing observations. To confront this challenge, we developed a Bayesian state space model for three species groups separately: (1) whole community, summed abundance of all species, (2) enhanced species, abundance of masu salmon, (3) unenhanced species, summed abundance of all species except masu salmon. A Bayesian state space model is best suited for our analysis because it can account for observation errors while imputing missing values given the long-term trend at each site. The model is composed of observation and state models as described below.

@@

$n_{s,t}$ is the "true" fish density excluding fish released in the spring, $Fry_{w(s),t}$ the number of salmon fry released in spring in watershed $w$ within which site $s$ is located, and $\beta_s$ the site-specific effect of released salmon fry on the observed fish density. The error term $\epsilon_{s,t}^{obs}$ is drawn from a normal distribution as $\epsilon_{s,t}^{obs} \sim Normal(0, \sigma_{obs,s}^2)$ to account for site- and year-specific observation errors. The parameter $\psi$ equals zero when we model the unenhanced species group (otherwise $\psi = 1$) so we exclude the term $\beta_s~Fry_{w(j),t}$.

@@

$$
\begin{aligned}
ln~n_{s,t+1} = ln~r_{s} + ln~n_{s,t} + \epsilon_{s,t}^{state} &\text{(8)}
\end{aligned}
$$ where $ln~r_{s}$ is the site-specific rate of change at site $s$ and $\epsilon_{s,t}^{state}$ is the process error that is assumed to follow a normal distribution as $\epsilon_{s,t}^{state} \sim Normal(0, \sigma_{state,s}^2)$. The site-specific rate of change is random draws from a normal distribution $ln~r_s \sim Normal(\mu_r, \sigma_r^2)$, assuming that community dynamics across Hokkaido have a shared temporal trend to some degree (the degree of shared trend is controlled by the standard deviation $\sigma_r$). This hierarchical structure allows for improved estimates of $ln~r_s$ by sharing information across sites. We used median estimates of fish density $n_{s,t}$ to calculate temporal CVs, means ($\mu$), and SDs ($\sigma$). Raw time series data and estimated community dynamics are shown in **Figures S12-S15**.

We developed the following linear regression model taking either CV, species richness, mean, or SD as a response variable $y_s$ with a normal or Poisson distribution.

@@

We fitted the models to the data using JAGS version 4.1.0 through *runjags* package in R. We assigned weakly informative priors for parameters, which were summarized in **Table S6**. Data manipulation and analysis were performed in R version 4.1.0.

>>>>

## raw data
source(here::here("code/data_fmt_fishdata.R"))

@@

## ssm estimate
df_bp <- list.files(path = here::here("data_fmt"), full.names = T) %>% 
  as_tibble() %>% 
  filter(str_detect(string = .$value, pattern = "ssm")) %>% 
  pull() %>% 
  lapply(function(x) {
    read_csv(x) %>% 
      filter(param == "bp_value") %>% 
      select(mean) %>% 
      mutate(group = str_extract(x, "all|masu|other"))
  }) %>% 
  bind_rows()

@@

**Affiliations:** ^1^ Department of Biology, University of North Carolina at Greensboro, Greensboro, NC 27412, USA ^2^ XXX ^3^ XXX ^4^ XXX

# Abstract

Although accumulating evidence suggests that stock enhancement with captive-bred organisms entails ecological risks, this method remains to be a common practice in conservation and fisheries. The widespread use of captive breeding is perhaps because of the "myth" that numerical benefits of released individuals may exceed the associated risks. Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of stock enhancement. Here, we use a simple theory to show that stock enhancement undermines long-term community stability by disrupting the sensitive balance of species interactions in natural communities. Stock enhancement destabilized community dynamics through facilitating competitive exclusion across a wide range of simulated scenarios. The present study further provided empirical support for this theoretical prediction using long-term data of Japanese stream fish communities. As our theory predicts, stream fish communities showed greater temporal fluctuations and reduced biodiversity in watersheds with intensive stocking of hatchery masu salmon - one of the most valuable freshwater resources in the region. Our findings paint a bleak picture for stock enhancement, reinforcing the recurrent calls for alternative management strategies.

@@

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient of species $j$ on species $i$, $K_i$ the carrying capacity, $R_t$ the number of released individuals for the enhancement of species 1, and $\epsilon_{i,t}$ the species response to stochastic environmental fluctuations that follow a normal distribution $Normal(0,\sigma_{\epsilon}^2)$. To capture variation in species traits, intrinsic growth rates of unenhanced species and interspecific competition coefficients were drawn randomly from a uniform ($r_{i,i \ne 1} \sim Unif(0,2)$) and an exponential distribution ($\alpha_{ij,i \ne j} \sim Exp(1/\bar{\alpha})$), respectively. The parameter $\phi_i$ controls the relative fitness of captive-bred individuals as follows:

@@

In our simulation, we considered `r nrow(param_set)` simulation scenarios with different values of the following ecological parameters: intrinsic growth of the enhanced species ($r_1=$ `r unique(sim_result$r1)`), average strength of interspecific competition ($\bar{\alpha}=$ `r unique(sim_result$alpha)`), carrying capacity ($K=$ `r unique(sim_result$k)`), degree of environmental stochasticity ($\sigma_{\epsilon}=$ `r unique(sim_result$sd_env)`), and relative fitness of captive-bred individuals ($f_R=$ `r unique(sim_result$phi)`). We chose these parameters because our sensitivity analysis revealed that they were influential in altering the relationship between community dynamics and stock enhancement (**Supporting Information**). Under each scenario with a unique parameter combination, we ran `r n_sim` time steps of `r max(sim_result$n_rep)` independent communities (i.e., simulation replicates) with varying numbers of released individuals $R$ ($0-500$). The last `r unique(sim_result$n_timestep)` time steps were used to obtain the following summary statistics of total community density $\sum_i^SN_i$: the coefficient of variation (CV), the number of species persist (defined as $N_i > 0.01$ at $t=1600$), the temporal mean ($\mu$), and the temporal SD ($\sigma$). In addition, the temporal mean and SD were also calculated for the enhanced ($N_1$) and unenhanced species ($\sum_{i,i \ne 1}^S N_i$) separately to infer mechanisms that underlie the community-level response.

@@

The only scenario where stock enhancement had little influence on community dynamics was when carrying capacity was sufficiently large ($K=500$; **Figures S7-S12**). In particular, stock enhancement increased the average density of the enhanced species with low population growth rate ($r_1=0.5$; **Figures S7 and S8**). This result may explain why some of the best evidence for successful stock enhancement comes from endangered species that typically suffer from increased mortality [@johnsonGeneticRestorationFlorida2010; @oppelPopulationReinforcementDemographic2021]. More importantly, the contrasting community response at different levels of carrying capacity casts light on the limitation of stock enhancement. One of the core motivations for stock enhancement is to mitigate or reverse declining trends of biodiversity due to human impacts, such as habitat loss and fragmentation [@chaseEcosystemDecayExacerbates2020]. Ironically, however, our results suggest that stock enhancement will not bring desired ecological outcomes unless we resolve root causes that compromise environmental capacity first.

@@

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of selective stock enhancement of masu salmon *Oncorhynchus masou masou* (**Figure 1E**) on the long-term stability of stream fish communities. Masu salmon is one of the most valuable freshwater resources in Hokkaido Island, Japan, and extensive hatchery programs have been implemented [@kitadaLessonsJapanMarine2020]. In protected watersheds (**Figure 1C and 1D**), long-term monitoring has been conducted over the past two decades along with official records of stock enhancement [@teruiMetapopulationStabilityBranching2018]. Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. In this study, we used the long-term data (1999-2019) of stream fish communities at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` protected watersheds (see Methods for selection criteria). Using hierarchical Bayesian models, we quantified the effect of stock enhancement on community dynamics while accounting for potential confounding effects of climates and local abiotic factors.

As our theory predicts, stream fish communities showed greater temporal fluctuations (higher CV) in watersheds with greater numbers of fish released (**Figure 1B**). The positive relationship between the CV and the number of fish released was strongly supported by the statistical model (**Figure 2**), in which the probability of the regression coefficient being positive was `r op(df_beta$all$prob_positive[df_beta$all$response == "cv"])` (**Table S8**). This pattern was associated with the reduced long-term average of total community density and fewer species richness (**Figure 1B and 2**), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figure 1B and 2;** see **Tables S8-S10** for full statistics). In the meantime, the SDs had vague relationships with stock enhancement, and neither positive nor negative relationship was supported by statistical analysis (**Figure 1B and 2**). At the population level, similar responses to stock enhancement have been reported in other fish and aquatic invertebrates [@arakiHatcheryStockingHelp2010]. Our analysis builds upon these findings by illustrating the community-level response.

@@

Like any non-experimental studies, we cannot exclude the possibility that our results were driven by spurious correlations. However, in the protected watersheds, any exploitation of aquatic species and alteration of in-stream habitats (e.g., construction of river-crossing structure) are strictly regulated; therefore, it is difficult to envision that unmeasured human influences played a major role in the observed relationships. Further, we have statistically controlled natural variation in environmental factors, including climatic and land use variables (see **Methods** and **Tables S8-S10**). This unique study setup may have helped uncover the close agreement between theory and empirical patterns.

@@

We employed a multispecies Ricker model [@fowlerSpeciesDynamicsAlter2012]. In the basic formula without stock enhancement, population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient between species $i$ and $j$, $K_i$ the carrying capacity, and $\epsilon_{i,t}$ the species response to stochastic environmental fluctuations that obey an normal distribution $Normal(0, \sigma_{\epsilon}^2)$. We modified this formula to include the effects of stock enhancement on reproduction and competition as follows:

@@

In this model, intrinsic growth rates of unenhanced species $r_{i,i \ne 1}$ and interspecific competition $\alpha_{ij}$ are random draws from a uniform ($r_{i,i \ne 1} \sim Unif(0,r_{max})$) and an exponential distribution ($\alpha_{ij} \sim Exp(1/\bar{\alpha})$), respectively. We assumed constant values of intraspecific competition ($\alpha_{ii}=1$), carrying capacity $(K_i = K)$ and the number of release $(R_t=R)$.

Prior to the main simulation analysis, we performed an extensive sensitivity analysis to identify parameters that strongly influence the relationship between community dynamics and stock enhancement (**Supporting Information**). We identified five influential parameters (**Table S2-4**), different values of which were considered in the main simulation as follows: intrinsic growth of enhanced species ($r_1 =$ `r unique(param_set$r1)`), average strength of interspecific competition ($\bar{\alpha}=$ `r unique(param_set$alpha)`), carrying capacity ($K=$ `r unique(param_set$k)`), degree of environmental stochasticity ($\sigma_{\epsilon}=$ `r unique(param_set$sd_env)`), and relative fitness of captive-bred individuals ($f_R=$ `r unique(param_set$phi)`). Meanwhile, we fixed values of the following parameters: the number of species ($S=$ `r unique(param_set$n_species)`) and the maximum intrinsic growth rate of unenhanced species ($r_{max}=$ `r unique(param_set$r_max)`). This simulation setup resulted in `r nrow(param_set)` sets of parameter combinations that cover a range of ecological scenarios.

Under each parameter combination, we ran `r n_sim` time steps of `r max(sim_result$n_rep)` independent communities (i.e., simulation replicates). The number of released individuals $R$ was drawn randomly from a uniform distribution for each simulation replicate as $R \sim Unif(0,500)$. We initialized the community with populations of each species drawn from a poisson distribution with a mean of five. We repeated the seeding procedure every 10 time steps over the first `r unique(sim_result$n_warmup)` time steps to allow species to establish populations and reach an equilibrium without stock enhancement (initialization period). After the initialization period, we released $R$ individuals of the enhanced species every time step over the next `r unique(sim_result$n_burnin)` time steps to reach a new equilibrium with selective stock enhancement (burn-in period). We continued the simulation run with stock enhancement and saved the last `r unique(sim_result$n_timestep)` time steps. We obtained the following summary statistics of the whole community $\sum_i^SN_j$: the coefficient of variation (CV), the number of species persist (defied as $N_i > 0.01$ at $t=1600$), the temporal mean ($\mu$), and the SD ($\sigma$). The temporal mean and SD were also calculated for the enhanced species ($N_1$) and unenhanced species ($\sum_{i,i \ne 1}^S N_i$) separately. We summarized values of simulation parameters in **Table S1**.

@@

**Time-series data.** We assembled time-series data of fish abundance across `r n_distinct(d0$site_id)` sites within `r n_distinct(d0$river)` protected watersheds of Hokkaido Island, Japan. At these sites, the long-term monitoring program began in 1971 but effective sampling methods (a combination of electrofishing and cast net sampling) were implemented in 1999. Most data were collected in summer to fall with irregular sampling intervals (0 to 3 year intervals for most cases). We confined our analysis to the sites that meets the following selection criteria: (1) the observation span (from the first to the last year of observation) exceeds 10 years, (2) the number of observation years exceeds five years, and (3) masu salmon are observed at least twice during the observation period. As a result, we used time-series data of fish abundance at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` watersheds during a period of 1999-2019. **Table S5** summarizes observed species in these watersheds.

**Fish stocking.** At the protected watersheds, stocking of masu salmon began in 1950s. A mixture of captive-bred and wild-born individuals were released in spring (fry [age 0+] and smolt [age 1+] stages) and fall (juvenile stage [age 0+]). For each watershed, we examined historical records and assembled annual data of stock enhancement (the number of fish released) for a period of 1999-2019. Although fish were released at multiple locations within a watershed, we were unable to extract stocking data at the site level.

**Environmental data.** At each sampling site, we measured the following environmental variables as potential covariates: upstream watershed area (km^2^; a proxy for stream size), proportional land use in the upstream watershed (forest), local climates (annual mean air temperature [degree Celsius] and cumulative precipitation [mm]), and ocean productivity (Chlorophyll a [unit]). We used MERIT Hydro [@yamazakiMERITHydroHighResolution2019] to delineate upstream watershed polygons for each sampling site. The proportion of forest land use in each watershed polygon was estimated based on land-use data in 2015 from Copernicus Global Land Service (100-m resolution) [@marcelbuchhornCopernicusGlobalLand2020]. Climate data at each sampling site were extracted from CHELSA ver. 1.2 [@kargerClimatologiesHighResolution2017; @kargerDataClimatologiesHigh2018]. Annual data of ocean productivity (2003-2019) were extracted from OceanColor (<https://oceancolor.gsfc.nasa.gov/>) and the averaged value within the 30-km radius of each river mouth was calculated. We used the following R packages to perform geospatial analysis: *sf* [@pebesmaSimpleFeaturesStandardized2018]*, raster* [@hijmansRasterGeographicData2020]*, exactextractr* [@bastonExactextractrFastExtraction2020]*, stars* [@pebesmaStarsSpatiotemporalArrays2020]*, whitebox* [@lindsayWhiteboxGATCase2016]*.*

@@

Our goal is to compare temporal community dynamics across sites. However, the data are not directly comparable because of observation errors (e.g., different observers) and missing observations. To confront this challenge, we developed a Bayesian state space model for three species groups separately: (1) whole community, summed abundance of all species, (2) enhanced species, abundance of masu salmon, (3) unenhanced species, summed abundance of all species except masu salmon. A Bayesian state space model is best suited for our analysis because it can account for observation errors while imputing missing values given the long-term trend at each site [@clarkPopulationTimeSeries2004a; @keryBayesianPopulationAnalysis2012]. The model is composed of observation and state models as described below.

@@

$n_{s,t}$ is the "true" fish density excluding fish released in the spring, $Fry_{w(s),t}$ the number of salmon fry released in spring in watershed $w$ within which site $s$ is located, and $\beta_s$ the site-specific effect of released salmon fry on the observed fish density. The parameter $\beta_s$ was drawn from a normal distribution with the hyper-mean $\mu_{\beta}$ and hyper-variance $\sigma_{\beta}^2$. The error term $\epsilon_{s,t}^{obs}$ is drawn from a normal distribution as $\epsilon_{s,t}^{obs} \sim Normal(0, \sigma_{obs,s}^2)$ to account for site- and year-specific observation errors. When modeling the unenhanced species group, $\psi$ equals zero (otherwise $\psi = 1$) so the model excludes the term $\beta_s~Fry_{w(s),t}$.

@@

$$
\begin{aligned}
ln~n_{s,t+1} = ln~r_{s} + ln~n_{s,t} + \epsilon_{s,t}^{state} &\text{(8)}
\end{aligned}
$$ where $ln~r_{s}$ is the site-specific rate of change at site $s$ and $\epsilon_{s,t}^{state}$ is the process error that is assumed to follow a normal distribution as $\epsilon_{s,t}^{state} \sim Normal(0, \sigma_{state,s}^2)$. The site-specific rate of change is random draws from a normal distribution $ln~r_s \sim Normal(\mu_r, \sigma_r^2)$, assuming that community dynamics across Hokkaido have a shared temporal trend to some degree (the degree of shared trend is controlled by the standard deviation $\sigma_r$). This hierarchical structure allows for improved estimates of $ln~r_s$ by sharing information across sites. We used median estimates of fish density $n_{s,t}$ to calculate temporal CVs, means ($\mu$), and SDs ($\sigma$). Bayesian p-values [@keryIntroductionWinBUGSEcologists2010] for the state-space models ranged from `r op(min(df_bp$mean))` to `r op(max(df_bp$mean))`, indicating that our model specification is appropriate. Estimated parameters and community dynamics are summarized in Supporting Information (**Table S7** and **Figures S13-S15**).

We developed the following linear regression model taking either CV, species richness (the number of species present during the observation period), mean, or SD as a response variable $y_s$ with a normal or Poisson distribution.

@@

We fitted the models to the data using JAGS version 4.1.0 through *runjags* package in R [@denwoodRunjagsPackageProviding2013]. We assigned weakly informative priors for parameters, which were summarized in **Table S6**. Data manipulation and analysis were performed in R version 4.1.0 [@rcoreteamLanguageEnvironmentStatistical2021].
